---
title: "class09_miniproject"
author: "Jocelyn Olvera"
date: "10/27/2021"
output:
  pdf_document: default
  html_document: default
---
# Class 9 Mini-Project
## 1. Exploratory data analysis

### Preparing the data

```{r, results='hide'}
# Save your input data file into your Project directory 
# (hid data with {r, results='hide'} to conserve pdf pages).
fna.data <- "WisconsinCancer.csv"
# Complete the following code to input the data and store as wisc.df
wisc.df <- read.csv(fna.data, row.names=1)
wisc.df
```



```{r, results='hide'}
# Examine the data (hid data with {r, results='hide'} to conserve pdf pages).
wisc.df
```



```{r}
# Examine the data with 'head()'
head(wisc.df)
```



```{r, results='hide'}
# We can use -1 here to remove the first column (hid data with {r, results='hide'} to conserve pdf pages).
wisc.data <- wisc.df[,-1]
wisc.data
```



```{r}
# Create diagnosis vector for later using 'as.factor()
diagnosis <- as.factor(wisc.df$diagnosis)
diagnosis
```



## Exploratory data analysis
### __Q1.__ How many observations are in this dataset?
#### There are 569 observations. 

```{r}
nrow(wisc.data)
```



#### __Q2.__ How many of the observations have a malignant diagnosis?
### 212 have a malignant diagnosis. 

```{r}
table(diagnosis)
```



#### class examples on retrieving column names. 

```{r}
colnames(wisc.df)
```



#### class example using 'grep' to help us find patterns we are interested in. 

```{r}
grep("_mean", colnames(wisc.df))
```


#### __Q3.__ How many variables/features in the data are suffixed with _mean?
### 10. 

```{r}
length(grep("_mean", colnames(wisc.df)))
```


# 2. Principal Component Analysis
## Performing PCA

```{r}
# Check column means and standard deviations
colMeans(wisc.data)
apply(wisc.data,2,sd)
```


#### __Q4.__ From your results, what proportion of the original variance is captured by the first principal components (PC1)?
### 0.4427

```{r}
# Perform PCA on wisc.data by completing the following code
wisc.pr <- prcomp(wisc.data, scale = TRUE)
```


```{r}
# Look at summary of results
summary(wisc.pr)
```

#### __Q5.__ How many principal components (PCs) are required to describe at least 70% of the original variance in the data?
### 3 are required. 

```{r}
# Summary of wisc.pr and assigning to new vector. 
x <- summary(wisc.pr)
x
```



```{r}
str(x)
```



```{r}
# Determining how many principal components (PCs) are required to describe at least 70% of the original variance in the data.
x$importance[3,] >= 0.7
PC70 <- which(x$importance[3,] >= 0.7)
PC70[1]
```

#### __Q6.__ How many principal components (PCs) are required to describe at least 90% of the original variance in the data?
### 3 are required.

```{r}
# Determining how many principal components (PCs) are required to describe at least 90% of the original variance in the data.
x$importance[3,] >= 0.9
PC90 <- which(x$importance[3,] >= 0.9)
PC90[1]
```

## Interpreting PCA results
#### Create a biplot of the wisc.pr using the biplot() function.


### __Q7.__ What stands out to you about this plot? Is it easy or difficult to understand? Why?
#### This plot is crowded and difficult to interpret.

```{r}
biplot(wisc.pr)
```

```{r, results='hide'}
# principal component (PC1/PC2)
# hid data with {r, results='hide'} to conserve pdf pages.
wisc.pr$x
```

```{r}
# Scatter plot observations by components 1 and 2
plot(wisc.pr$x, col = diagnosis , 
     xlab = "PC1", ylab = "PC2")
```

### __Q8.__ Generate a similar plot for principal components 1 and 3. What do you notice about these plots?
## Although similar, PC1 vs PC2 displays less overlap between the data sets.

```{r}
# Repeat for components 1 and 3
plot(wisc.pr$x[, c(1, 3)], col = diagnosis, 
     xlab = "PC1", ylab = "PC3")
```

## ggplot

```{r}
# Create a data.frame for ggplot
df <- as.data.frame(wisc.pr$x)
df$diagnosis <- diagnosis

# Load the ggplot2 package
library(ggplot2)

# Make a scatter plot colored by diagnosis
ggplot(df) + 
  aes(PC1, PC2, col=diagnosis) + 
  geom_point()
```


## Variance explained

```{r}
# Calculate variance of each component
pr.var <- wisc.pr$sdev^2
head(pr.var)
```


```{r}
sum.pr <- sum(pr.var)
```


#### Calculate the variance explained by each principal component by dividing by the total variance explained of all principal components. Assign this to a variable called pve and create a plot of variance explained for each principal component.

```{r}
# Variance explained by each principal component: pve
pve <- pr.var/ sum.pr
pve

# Plot variance explained for each principal component
plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")
```


```{r}
# Alternative scree plot of the same data, note data driven y-axis
barplot(pve, ylab = "Precent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )
```


```{r}
## ggplot based graph
#install.packages("factoextra")
library(factoextra)
fviz_eig(wisc.pr, addlabels = TRUE)
```


## Communicating PCA results


### __Q9.__ For the first principal component, what is the component of the loading vector (i.e. wisc.pr$rotation[,1]) for the feature concave.points_mean?
#### concave.points_mean = -0.26085376 

```{r}
wisc.pr$rotation[,1]
```

### __Q10.__ What is the minimum number of principal components required to explain 80% of the variance of the data?
#### Minimum number of principal components is 5. 

```{r}
x$importance[3,] >= 0.8
PC80 <- which(x$importance[3,] >= 0.8)
PC80[1]
```


# 3. Hierarchical clustering
### First scale the wisc.data data and assign the result to data.scaled.
```{r}
# Scale the wisc.data data using the "scale()" function
data.scaled <- scale(wisc.data)
```


### Calculate the (Euclidean) distances between all pairs of observations in the new scaled dataset and assign the result to data.dist.
```{r}
data.dist <- dist(data.scaled)
```


### Create a hierarchical clustering model using complete linkage. Manually specify the method argument to hclust() and assign the results to wisc.hclust.

```{r}
wisc.hclust <- hclust(data.dist, method="complete")
```


## Results of hierarchical clustering

### Q11. Using the plot() and abline() functions, what is the height at which the clustering model has 4 clusters?
#### height = 19

```{r}
plot(wisc.hclust)
abline(h=19, col="red", lty=2)
```

## Selecting number of clusters
### Use cutree() to cut the tree so that it has 4 clusters. Assign the output to the variable wisc.hclust.clusters.


```{r}
wisc.hclust.clusters <- cutree(wisc.hclust, k=4)
table(wisc.hclust.clusters, diagnosis)

```


### __Q12.__ Can you find a better cluster vs diagnoses match by cutting into a different number of clusters between 2 and 10?

```{r}
wisc.hclust.clusters2 <- cutree(wisc.hclust, k = 2, 10)
table(wisc.hclust.clusters2, diagnosis)
```

### __Q13.__ Which method gives your favorite results for the same data.dist dataset? Explain your reasoning.
#### The 'ward.D2' method appears to give a better result of the same data.dist data set. The dendogram displays the clusters generated from squaring the dissimilarites. This gives a cleaner dendogram. 

```{r}
wisc.pr.hclust <- hclust(data.dist, method="ward.D2")
plot(wisc.pr.hclust)
```

```{r}
wisc.pr.hclust.C <- hclust(data.dist, method="complete")
plot(wisc.pr.hclust.C)
```




# 4. OPTIONAL: K-means clustering

```{r}
wisc.km <- kmeans(scale(wisc.data), centers= 2, nstart= 20)
table(wisc.km$cluster, diagnosis)
```

### __Q14.__ How well does k-means separate the two diagnoses? How does it compare to your hclust results?
#### Fewer false positives/negatives.


```{r}
table(wisc.hclust.clusters, wisc.km$cluster)
```


# 5. Combining methods
## Clustering on PCA results

```{r}
grps <- cutree(wisc.pr.hclust, k=2)
table(grps)
```

```{r}
# We can do a cross-table by giving the table() function two inputs. 
# TP (true positive = 164) FP (false positive = 20)
# TN (true negatives = 337) FP (false positive = 48)
table(grps, diagnosis)
```


```{r}
plot(wisc.pr$x[,1:2], col=grps)
```

```{r}
g <- as.factor(grps)
levels(g)
```

```{r}
g <- relevel(g,2)
levels(g)
```

```{r}
# Plot using our re-ordered factor 
plot(wisc.pr$x[,1:2], col=g)
```

## OPTIONAL

```{r}
library(rgl)
plot3d(wisc.pr$x[,1:3], xlab="PC 1", ylab="PC 2", zlab="PC 3", cex=1.5, size=1, type="s", col=grps)
```


```{r}
## Use the distance along the first 7 PCs for clustering i.e. wisc.pr$x[, 1:7]
wisc.dist <- dist(wisc.pr$x[, 1:7])
wisc.pr.hclust <- hclust(wisc.dist, method="ward.D2")
plot(wisc.pr.hclust)
```

### __Q15.__ How well does the newly created model with four clusters separate out the two diagnoses?
```{r}
wisc.pr.hclust.clusters <- cutree(wisc.pr.hclust, k=2)
table(wisc.pr.hclust.clusters, diagnosis)
```


### __Q16.__ How well do the k-means and hierarchical clustering models you created in previous sections (i.e. before PCA) do in terms of separating the diagnoses? Again, use the table() function to compare the output of each model (wisc.km$cluster and wisc.hclust.clusters) with the vector containing the actual diagnoses.

```{r}
table(wisc.km$cluster, diagnosis)
```

```{r}
table(wisc.hclust.clusters, diagnosis)
```


#6. Sensitivity/Specificity
### __Q17.__ Which of your analysis procedures resulted in a clustering model with the best specificity? How about sensitivity?
#### It appears that the best specificity comes from using 'hclust' and we get the best sensitivity with 'kmeans'. 

#### calculating sensitivity TP/(TP+FN)
```{r}
(175)/(175+14)
```


# 7. Prediction

```{r}
#url <- "new_samples.csv"
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```

```{r}
plot(wisc.pr$x[,1:2], col=g)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")
```


### __Q18.__ Which of these new patients should we prioritize for follow up based on your results?
#### patient 1. 
