---
title: "class18_genome"
author: "Jocelyn Olvera"
date: "12/1/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r}
library(ggplot2)
library(dplyr)
```


```{r}
expr <- read.table("../class18_genome/rs8067378_ENSG00000172057.6.txt")
head(expr)
```

```{r}
nrow(expr)
```

```{r}
table(expr$geno)
```

```{r}
summary(expr)
```


## __Q13:__ Read this file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes. Hint: The read.table(), summary() and boxplot() functions will likely be useful here. There is an example R script online to be used ONLY if you are struggling in vein.  Note that you can find the medium value from saving the output of the boxplot() function to an R object and examining this object. There is also the medium() and summary() function that you can use to check your understanding.

### __Response:__ 
### A/A median = 31.24847
### A/G median = 25.06486
### G/G median = 20.07363


```{r}
ggplot(expr,
       aes(x = geno,
           y = exp)) +
  geom_boxplot()  
  
  
expr %>%
 group_by(geno) %>%
  summarise(Min = min(exp),
            Max = max(exp),
            Median = median(exp))
```

## __Q14:__ Generate a boxplot with a box per genotype, what could you infer from the relative expression value between A/A and G/G displayed in this plot? Does the SNP effect the expression of ORMDL3? Hint: An example boxplot is provided overleaf – yours does not need to be as polished as this one.
### __Response:__ A/A is more highly expressed than G/G in this gene. However, the presence of G/G is associated with a decreased expression of this gene. therefore, the SNP appears to effect the expression of the ORMDL3 gene and is inferred by the presence of G/G. 


```{r}
bp <- ggplot(expr) + aes(geno, exp, fill=geno) +
  labs(y = "expression", x = "genotype") +
  geom_boxplot(notch = TRUE)
bp

```






## Questions #5 and 6

#### Data downloaded from: https://uswest.ensembl.org/Homo_sapiens/Variation/Sample?db=core;r=17:39903525-39926526;v=rs8067378;vdb=variation;vf=105535077#373531_tablePanel



## Section1: Proportion of G/G in a population
```{r}
mxl <- read.csv("../class18_genome/373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(mxl)
```


```{r}
table(mxl$Genotype..forward.strand.)
```


# Total number of visuals in the dataset. Percentage of homozygous.
```{r}
table(mxl$Genotype..forward.strand.) / nrow(mxl) *100
```


```{r}
gbr <- read.csv("../class18_genome/373522-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
```



## Find proportion of GIG
```{r}
round(table(gbr$Genotype..forward.strand.) / nrow(gbr) *100, 2)
```


## This variant that is associated with childhood asthama is more frequent in the GBR population thatn the MKL population. 

##Lets now dig into this further. 


