---
title: "class17_vaccines"
author: "Jocelyn Olvera"
date: "11/24/2021"
output: 
  pdf_document:
    latex_engine: xelatex
---


#Looking at COVID-19 vaccination rates during the Holiday travel 
#Data obtained from: https://data.ca.gov/dataset/covid-19-vaccine-progress-dashboard-data-by-zip-code


### Importing Data
```{r}
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
head(vax)
```

## Q: How many of entries 'nrow(vax)'? 
```{r}
nrow(vax)
```

## Using **skimr** package and the 'skim()' function to get a quick overview of the data. 
### using 'skimr::skim' allows you to use the package only once without loading it in with 'library()'.
```{r}
skimr::skim(vax)
```




## Working with dates
Notice that one of these columns is a date column. Working with time and dates gets annoying quickly. We can use **lubridate** package to make this easy...

```{r}
library(lubridate)
today()
```

## Q1. What column details the total number of people fully vaccinated? 
## Column: "persons_fully_vaccinated"

## Q2. What column details the Zip code tabulation area? 
## Column: "zip_code_tabulation_area"

## Q3. What is the earliest date in this dataset? 
```{r}
vax$as_of_date[1]
```

##This will not work because our data column was read as character...
```{r}
# today()- vax$as_of_date[1]
```

##Instead, try overwritting the first column to match the format and assign to 'd' vector. 
```{r}
d <- ymd(vax$as_of_date[1])
```

```{r}
today() - d[1]
```

##I will make the 'as_of_date' column Date format... 
```{r}
vax$as_of_date <- ymd(vax$as_of_date)
```


##Q. When was the dataset last updated?
# Q4. What is the latest date in this dataset? 

```{r}
vax$as_of_date[nrow(vax)]
```
## Q9. How many days have passed since the last update of the dataset? 

```{r}
today() - vax$as_of_date[nrow(vax)]
```

##Q. How many days does the dataset span?

```{r}
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```

## Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)? 
##Q. How many different ZIP code areas are in this dataset?

```{r}
length(unique(vax$zip_code_tabulation_area))
```




## Working with ZIP codes
##To work with ZIP codes we use the **zipcodeR** 

```{r}
library(zipcodeR)
```

```{r}
geocode_zip('92037')
```


## Focus on the San Diego area
```{r}
inds <- vax$county == "San Diego"
nrow(vax[inds,])
```

## Sub-setting can get tedious and complicated quickly when you have multiple things we want to subset by. 
```{r}
library(dplyr)
```

## We will use the 'filter()' function to do our subsetting from now on. 
## We want to focus on San Diego. 

```{r}
sd <- filter(vax, county == "San Diego")
nrow(sd)
```

## More subsetting. 
```{r}
sd.20 <- filter(vax, county == "San Diego" &
                age5_plus_population > 20000)
nrow(sd.20)
```


## Q11. How many distinct zip codes are listed for San Diego County? 
```{r}
length(unique(sd))
```

## Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset?

## Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2021-11-09”? 
```{r}
sd.now <- filter(vax, county == "San Diego",
                as_of_date == "2021-11-23")
                
head(sd.now)
```

```{r}
sd.now$percent_of_population_fully_vaccinated
```


```{r}
sd.vax <- sd.now$percent_of_population_fully_vaccinated
summary(sd.vax)
```


## Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2021-11-09”?


Base R histogram
```{r}
hist(sd.now$percent_of_population_fully_vaccinated)
```
### This plot above is going to be susceptible to being skewed byt ZIP code areas with small populations. This will have big effects for just a small number of unvaxed folks...

##Q. What is the population population of the 92037 ZIP code area?

```{r}
sd.92037 <- filter(sd.now, zip_code_tabulation_area == "92037")
head(sd.92037)
```

##Q. What is the average vaccination value for this UCSD/La JOlla ZIP code area? 

```{r}
sd.92037$age5_plus_population
```

```{r}
sd.92037$percent_of_population_fully_vaccinated
```

## Area interested in, Escondido, CA 92025.
```{r}
sd.92025 <- filter(vax, zip_code_tabulation_area == "92025")
sd.92025$age5_plus_population
sd.92025$percent_of_population_fully_vaccinated
```

## Area interested in by classmate 92124.
```{r}
sd.92124 <- filter(sd.now, zip_code_tabulation_area == "92124")
sd.92124$age5_plus_population
sd.92124$percent_of_population_fully_vaccinated
```


```{r}
sd.92103 <- filter(sd.now, zip_code_tabulation_area == "92103")
sd.92103$age5_plus_population
sd.92103$percent_of_population_fully_vaccinated
```


## Focus on UCSD/La Jolla

## Plotting area of interest, begining with La Jolla (92037). 
```{r}
sd.92037 <- filter(vax, zip_code_tabulation_area == "92037")
```

```{r}
library(ggplot2)
ggplot(sd.92037) +
  aes(x = as_of_date,
      y = percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x= "Date", y="Percent Vaccinated")
```


## For Escondido, CA 92025.
```{r}
library(ggplot2)
ggplot(sd.92025) +
  aes(x = as_of_date,
      y = percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x= "Date", y="Percent Vaccinated")
```

## Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-16”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?




## Lets make this plot for all San Diego County ZIP code areas that have a population as least as large as 92037. 

```{r}
head(vax$age5_plus_population)
```

```{r}
sd.36 <- filter(vax, county == "San Diego",
                 age5_plus_population > 36144)
head(sd.36)
```

## How many ZIP code areas in San Diego County have a population larger than 92037. 

```{r}
length(unique(sd.36$zip_code_tabulation_area))
```

```{r}
library(ggplot2)
ggplot(sd.36) +
  aes(x = as_of_date,
      y = percent_of_population_fully_vaccinated,
      group = zip_code_tabulation_area) +
  geom_line()
  labs(x= "Date", y="Percent Vaccinated")
```


## Making a plot for all California with populations as large as La Jolla. 
```{r}
ca.all <- filter(vax,
                 age5_plus_population > 36144)
head(ca.all)
```

## How many ZIP codes in CA are as large as La Jolla. 
```{r}
length(unique(ca.all$zip_code_tabulation_area))
```


## Make the plot for all the ZIP codes as large as La Jolla. 
```{r}
library(ggplot2)
ggplot(ca.all) +
  aes(x = as_of_date,
      y = percent_of_population_fully_vaccinated,
      group = zip_code_tabulation_area) +
  geom_line(alpha = 0.2) +
  labs(x= "Date", y="Percent Vaccinated") +
  geom_hline(yintercept=0.67, color = "red")
```



## What is the mean across the state for these 36k + population areas. 

```{r}
ca.now <- filter(ca.all,
                 as_of_date == "2021-11-23")
summary(ca.now$percent_of_population_fully_vaccinated)
```
